---
title: "Baby-Names"
author: "Mia Fitzgerald"
date: "4/2/2026"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(knitr)
library(tidyverse)
library(babynames)
library(janitor)
```


### Mutated Proportion Into Rounded Percentages 


```{r}
data(babynames)

babynames2 <- babynames %>%
  mutate(prop_percent = round(prop * 100, 2))

babynames2 %>% 
  head(6) %>%
  kable()

```


I load the babynames dataset and use mutate() to create a new column called prop_percent by multiplying the original prop by 100 and rounding the result to two decimal places. I then take the first six rows of this updated data and pass them into kable() to display a clean, formatted table.
^[Question 1 complete]


### Using Strings in a New Column 


```{r}
babynames3 <- babynames2 %>%
  mutate(percent_year = str_glue("{prop_percent}% in {year}"))

babynames3 %>% 
  head(6) %>%
  kable()

```


I take the babynames2 dataset and use mutate() with the str_glue() function to create a new column called percent_year that combines the prop_percent and year values into a single formatted string. I then select the first six rows of this updated data and pass them into kable() to display a table. 
^[Question 2 complete]


### Filtered For Specific Name and Sex


```{r}
babynames4 <- babynames3 %>%
  filter(name == "Cora", sex == "F")

babynames4 %>%
  head(6) %>%
  kable()

```


I take the babynames3 dataset and use filter() to create a new object called babynames4 that only includes rows where the name is "Cora" and the sex is "F". I then take the first six rows of this filtered data and pass them into kable() to display the results in a table.
^[Question 3 complete]


### Five Random Rows Using set.seed()


```{r}
set.seed(42)
babynames_sliced <- babynames4 %>%
  slice_sample(n = 5) %>%
  arrange(year)

babynames_sliced %>%
  kable()

```


I start by using set.seed(42) and then take the babynames4 dataset and use slice_sample(n = 5) to randomly select five rows, followed by arrange(year) to sort those rows in chronological order. Finally, I pass this sampled and sorted data into kable() to display the results in a table.
^[Question 4 complete]


### Making a Collapsed String


```{r}
babystring <- babynames_sliced %>%
  pull(percent_year) %>%
  str_flatten(collapse = ", ", last = ", and ")

babystring %>%
  str_view()

```


I take the babynames_sliced dataset and use pull() to extract the percent_year column. I use str_flatten() to merge those values into a single string, separating them with commas and adding ", and ". Finally, I pass this combined string into str_view() to display the resulting text in a formatted view.
^[Question 5 complete]


### Gluing Strings  


```{r, results= 'asis'}
cat(str_glue("Out of all female names, the name Cora represented {babystring}."))

```


I use str_glue() to create a sentence that embeds babystring into a specific text string about the name Cora. I then use cat() to print this final sentence, while the results='asis' chunk option ensures that the text is rendered as standard Markdown in my document.
^[Question 6 complete]